Automated Verification and Synthesis of Embedded Systems: Advances and Challenges

### 2017-06-21 paper.tex
minor changes to new text

As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program in single static assignment (SSA) form~\cite{Appel97} is shown in Figure~\ref{figure:verification-condition}(b). Note that the SSA form is an intermediate representation used by compilers to facilitate optimizations and transformations of the program code. The common property in SSA form is that every variable has only one definition in the program text. This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript) such that every assignment has a unique left hand side as shown in Figure~\ref{figure:verification-condition}(b).
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int x[N], a;
...
int main(void) {
  a=N;
  for(int i=0; i<N; i++)
    if (x[i]>1)
      a--;
  assert(a<=N);
  return 0;
}
\end{lstlisting}
\end{minipage}
\begin{center}
(a)
\end{center}
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
a1 = N
a2 = (x[0] > 1) ? a1 - 1 : a1
a3 = (x[1] > 1) ? a2 - 1 : a2
a4 = (x[2] > 1) ? a3 - 1 : a3
...
an+1 = (x[N-1] > 1) ? an - 1 : an 
\end{lstlisting}
\end{minipage}
\begin{center}
(b)
\end{center}
\caption{(a) A simple C program with a \textit{for} loop. (b) The corresponding unwound C program of (a) converted into SSA form.} 
\label{figure:verification-condition}
\end{figure}

Apart from that, this program has an exponential number of paths since each element of array $x$ can be either greater than one or less than or equal to one. Despite the large number of paths through the program, BMC unwinds it up to a bound $k$ and translates it into a VC $\psi$ such that $\psi$ is satisfiable if and only if the assertion $\left(a<=N\right)$ fails. Note that BMC still encodes the states of the program with a size that grows linearly with N. More precisely, the program in Figure~\ref{figure:verification-condition}(a) is converted into the $\psi$ using a decidable fragment of first-order logic as follows~\cite{Bradley07}:
%
\begin{equation}              
\label{verification-condition-example}
\psi := \left [ \begin{array}{ll} 
                a_{1} = N  \\
                \wedge \, \, a_{2} = ite\left(x[0] > 1, a_{1} - 1, a_{1}\right) \\ 
                \wedge \, \, a_{3} = ite\left(x[1] > 1, a_{2} - 1, a_{2}\right) \\
                \wedge \, \, a_{4} = ite\left(x[2] > 1, a_{3} - 1, a_{3}\right) \\
                \wedge \ldots \\
                \wedge \, \, a_{N+1} = ite\left(x[N-1] > 1, a_{N} - 1 , a_{N}\right) \\
                \wedge \neg \left(a_{N+1} \leq N\right) \\
              \end{array} \right ]  \\
              \\           
\end{equation}

The ternary operator $f \: \: ? \: \: t_1 : t_2$ shown in Figure~\ref{figure:verification-condition}(b) is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and depending on its value selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$). In order to verify that the assertion $\left(a<=N\right)$ holds, its negation is added to $\psi$ and we check whether the entire formula is satisfiable using an off-the-self SMT solver. Formula (\ref{verification-condition-example}) can simply be represented as a Boolean logic circuit, which can further be transformed into a (equisatisfiable) CNF formula over propositional variables by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables. However, checking the validity of a first-order logic formula in a given background theory is an ${\cal NP}$-complete problem~\cite{PatarinG97}).

>>>>

As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b). One may notice that the SSA form is an intermediate representation, used by compilers, to facilitate optimizations and transformations of source code. The common property in SSA form is that every variable has only one definition in the program text. This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left hand side, as shown in Figure~\ref{figure:verification-condition}(b).
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int x[N], a;
...
int main(void) {
  a=N;
  for(int i=0; i<N; i++)
    if (x[i]>1)
      a--;
  assert(a<=N);
  return 0;
}
\end{lstlisting}
\end{minipage}
\begin{center}
(a)
\end{center}
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
a1 = N
a2 = (x[0] > 1) ? a1 - 1 : a1
a3 = (x[1] > 1) ? a2 - 1 : a2
a4 = (x[2] > 1) ? a3 - 1 : a3
...
an+1 = (x[N-1] > 1) ? an - 1 : an 
\end{lstlisting}
\end{minipage}
\begin{center}
(b)
\end{center}
\caption{(a) A simple C program with a loop \textit{for}. (b) The corresponding unwound C program of (a) converted into SSA form.} 
\label{figure:verification-condition}
\end{figure}

Apart from that, this program has an exponential number of paths, since each element of $x$ can be either greater than one or less than or equal to one. Despite the large number of paths through that program, BMC unwinds it up to a bound $k$ and translates it into a VC $\psi$, such that $\psi$ is satisfiable if and only if the assertion $\left(a<=N\right)$ fails. One may also notice that BMC still encodes program stages with a size that grows linearly with $N$. More precisely, the program in Figure~\ref{figure:verification-condition}(a) is converted into the $\psi$ using a decidable fragment of first-order logic, as follows~\cite{Bradley07}:
%
\begin{equation}              
\label{verification-condition-example}
\psi := \left [ \begin{array}{ll} 
                a_{1} = N  \\
                \wedge \, \, a_{2} = ite\left(x[0] > 1, a_{1} - 1, a_{1}\right) \\ 
                \wedge \, \, a_{3} = ite\left(x[1] > 1, a_{2} - 1, a_{2}\right) \\
                \wedge \, \, a_{4} = ite\left(x[2] > 1, a_{3} - 1, a_{3}\right) \\
                \wedge \ldots \\
                \wedge \, \, a_{N+1} = ite\left(x[N-1] > 1, a_{N} - 1 , a_{N}\right) \\
                \wedge \neg \left(a_{N+1} \leq N\right) \\
              \end{array} \right ]  \\
              \\           
\end{equation}

The ternary operator $f \: \: ? \: \: t_1 : t_2$, shown in Figure~\ref{figure:verification-condition}(b), is converted into the conditional expression $\mathit{ite}(f, t_1, t_2)$ that takes as its first argument the Boolean formula $f$ and, depending on its value, selects either the second (i.e., $t_1$) or the third argument (i.e., $t_2$). In order to verify that the assertion $\left(a<=N\right)$ holds, its negation is added to $\psi$ and it is checked whether the entire formula is satisfiable, through an off-the-self SMT solver. Formula (\ref{verification-condition-example}) can be represented simply as a Boolean logic circuit, which can further be transformed into a (equisatisfiable) CNF formula over propositional variables, by Tseitin's transform~\cite{Tseitin83} in linear time and by introducing at most a linear number of fresh variables; however, checking the validity of a first-order logic formula, in a given background theory, is an ${\cal NP}$-complete problem~\cite{PatarinG97}).
